---
title: "Biobank project"
output: html_notebook
---

Read data into R
```{r}
setwd("~/IDrive-Sync/Biobank/data/")
library(data.table)
bd<-fread(file="ukb49753.tab", sep='\t', header=T, stringsAsFactors = F)
#bd<-fread(file="ukb49753.tab", sep='\t', header=T, quote="", stringsAsFactors = F) #In response to: Found and resolved improper quoting out-of-sample. First healed line 84910. If the fields are not quoted (e.g. field separator does not appear within any field), try quote="" to avoid this warning.
```

Can generate counts for each category by running this code (will generate tables of results too):

```{r}
happy.hated<-table(bd$f.20487.0.0, bd$f.20458.0.0) #happy v hated as a child
write.csv(happy.hated, "happy-hated_as_child.csv")

happy.loved<-table(bd$f.20489.0.0, bd$f.20458.0.0) #happy v loved as a child
write.csv(happy.loved, "happy-loved_as_child.csv")

happy.phyAbus<-table(bd$f.20488.0.0, bd$f.20458.0.0) #happy v physically abused as child
write.csv(happy.phyAbus, "happy-phys_abuse_as_child.csv")

happy.sexAbus<-table(bd$f.20490.0.0, bd$f.20458.0.0) #happy v sexually abused as a child
write.csv(happy.sexAbus, "happy-sex_abuse_as_child.csv")

miser.hated<-table(bd$f.20487.0.0, bd$f.1930.0.0) #miserable v hated as a child
write.csv(miser.hated, "miser-hated_as_child.csv")

miser.loved<-table(bd$f.20489.0.0, bd$f.1930.0.0) #miserable v loved as a child
write.csv(miser.loved, "miser-loved_as_child.csv")

miser.phyAbus<-table(bd$f.20488.0.0, bd$f.1930.0.0) #miserable v physically abused as child
write.csv(miser.phyAbus, "miserable-physical_abuse_as_child.csv")

miser.sexAbus<-table(bd$f.20490.0.0, bd$f.1930.0.0) #miserable v sexually abused as a child
write.csv(miser.sexAbus, "miserable-sex_abuse_as_child.csv")

miser.happy<-table(bd$f.1930.0.0, bd$f.20458.0.0) #miserable le v happy
write.csv(miser.happy, "miserable_happy.csv")

hated.loved<-table(bd$f.20487.0.0, bd$f.20489.0.0) #felt loved v felt hated
write.csv(hated.loved, "hated_loved_as_child.csv")
```

To further explore patterns, generate new columns that earmark whether the person endorses 'often' or 'very often' for hated as a child, physically abused and sexually abused.


Then use these new columns to figure out how many have all 3, how many have 2 or more and how many have 1 or more


Also generate a variable that indicates "happy" for 1,2,3 and unhappy for 4,5,6


TRAUMA: "Physically and sexually molested", "Physically or sexually molested", "Neither" (never = 0, rarely, sometimes, often, very often = 1). So for each person look at their 'phyAbus' and 'sexAbus' columns

```{r}
bd<-bd %>% 
	select(name, 20488) %>%
	mutate(phyAbus = case_when(
			20488=="2" ~ "1",
			20488=="3" ~ "1",
			20488=="4" ~ "1",
			20488=="0" ~ "0",
			20488=="1" ~ "1",
			20488=="-818" ~ NA_character_

))
```



```{r}
bd<-bd %>% 
	select(name, 20490) %>%
	mutate(sexAbus = case_when(
			20490=="2" ~ "1",
			20490=="3" ~ "1",
			20490=="4" ~ "1",
			20490=="0" ~ "0",
			20490=="1" ~ "1",
			20490=="-818" ~ NA_character_

))
```



```{r}
bd <- bd %>%
  mutate(trauma=case_when(
    physAbus=="1" & sexAbus=="1" ~ "2",
    physAbus=="1" & sexAbus=="0" ~ "1",
    physAbus=="0" & sexAbus=="1" ~ "1",
    physAbus=="0" & sexAbus=="0" ~ "0",
    physAbus=="0" & is.na(sexAbus) ~ "0",
    physAbus=="1" & is.na(sexAbus) ~ "1",
    is.na(physAbus) & sexAbus=="1" ~ "1",
    is.na(physAbus) & sexAbus=="0" ~ "0",
    is.na(physAbus) & is.na(sexAbus) ~ "0"
  ))
```


Then do table of TRAMUA v happiness


Just as an aside, you can look at whether childhood adversity predicts adult adversity

